---
output: github_document
---

## Statuses

[![Travis-CI Build Status](https://travis-ci.org/fboehm/gemma2.svg?branch=master)](https://travis-ci.org/fboehm/gemma2)

[![codecov](https://codecov.io/gh/fboehm/gemma2/branch/master/graph/badge.svg)](https://codecov.io/gh/fboehm/gemma2)



<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview

`gemma2` is an implementation in R of the GEMMAv 0.97 EM algorithm that is part of the GEMMA algorithm for REML estimation of multivariate linear mixed effects models of the form:

$vec(Y) = X vec(B) + vec(G) + vec(E)$ where $E$ is a n by 2 matrix of random effects that follows the matrix-variate normal distribution

$G \sim MN(0, K, V_g)$ where $K$ is a relatedness matrix and $V_g$ is a 2 by 2 covariance matrix for the two traits of interest.

Additionally, the random errors matrix $E$ follows the distribution:

$E \sim MN(0, I_n, V_e)$

and $G$ and $E$ are independent.





## Installation

To install `gemma2`, use the `devtools` R package from CRAN. If you haven't installed `devtools`, please run this line fo code:

```{r, eval = FALSE}
install.packages("devtools")
```

Then, run this line of code to install `gemma2`:

```{r, eval = FALSE}
devtools::install_github("fboehm/gemma2")
```



